Evolution of pre-copulatory and post-copulatory inbreeding behaviour and associated

### 2016-12-16 ms_pre_post.Rmd
Edits up through end of Model

Methods
===================================================

@@

@@

To test the hypothesis that evolution of post-copulatory inbreeding avoidance can facilitate evolution of costly polyandry, we compare simulations in which polyandry and pre-copulatory and post-copulatory inbreeding can all evolve, with otherwise identical simulations in which post-copulatory inbreeding strategy cannot evolve given four different costs of polyandry ($c_{P} = \{0, 0.0025, 0.005,  0.01\}$). To prevent $F_{p}$ from evolving, we sever the connection from $F_{a}$ to $F_{p}$ such that all $F_{g}$ genotypes cause random fertilisation with respect to kinship. Hence $F_{a}$ alleles have no phenotypic effect, meaning that polyandrous females cannot enact post-copulatory inbreeding avoidance.

To test the hypothesis that selection for a costly strategy of inbreeding avoidance will be negligible given a less costly alternative strategy, we quantify change in $M_{a}$ and $F_{a}$ over generations in simulations where the cost of pre-copulatory inbreeding strategy $c_{M}=0$ but the cost of post-copulatory inbreeding strategy $c_{F}=0.02$, and vice versa. We compare these simulations with evolution of a costly strategy in the absence of evolution of an alternative strategy (e.g., evolution of pre-copulatory inbreeding strategy when post-copulatory inbreeding strategy phenotype is fixed at zero, $F_{p}=0$).

To test the hypothesis that the existence of adaptive post-copulatory inbreeding avoidance will preclude the invasion of pre-copulatory inbreeding avoidance, or vice versa, we first used exploratory simulations to determine evolution of pre-copulatory inbreeding strategy, and post-copulatory inbreeding strategy and associated polyandry, in isolation. To test whether pre-copulatory inbreeding avoidance would evolve when adaptive polyandry and post-copulatory inbreeding avoidance was already established, we initiated $M_{a}$ allele values at zero, but fixed $F_{a}$ allele values at $-10$ and $P_{a}$ allele values at $1$ (i.e., meaning that $F_{p}$ and $P_{p}$ did not evolve). Similarly, to test whether post-copulatory inbreeding avoidance evolved when pre-copulatory inbreeding avoidance was already established, we initiated $F_{a}$ and $P_{a}$ allele values at zero but fixed $M_{a}$ allele values at $-10$. Consequently, because females have 10 diploid loci, when $M_{a}$ or $F_{a}$ alleles were fixed at $-10$, outbred females were $51$ times less likely to choose a full brother and $13.5$ times times less likely to choose a first cousin than a non-relative in pre-copulatory and post-copulatory choice, respectively.

In all simulations, we record mean values of $P_{a}$, $M_{a}$, $F_{a}$ in each generation. Each combination of parameter values simulated is replicated $40$ times, and grand mean values and standard errors of means are calculated in each generation across replicates. These analyses allow us to infer how allele values change over generations in response to costs, but also in response to the changing values of other alleles and therefore potentially interacting phenotypes (e.g., selection on post-copulatory inbreeding strategy might be weak if polyandry is rare). For all replicates, we set the maximum number of generations to 40000, which exploratory simulations and previous modelling [@Duthie] showed to be a sufficient number of generations for inferring long-term dynamics of mean allele values and therefore selection on phenotypes.  For all replicates, we set $\rho=5$ immigrants, which produced a range of kin and non-kin in each generation allowing females to express inbreeding strategies, and $n=8$ offspring, which was sufficient to keep populations consistently at carrying capacities and avoid population extinction. Values of $K_{f}$ and $K_{m}$ were both set to 100 because previous modelling has shown that populations of this size are small enough that mate encounters between kin occur with sufficient frequency for selection on inbreeding strategy, but populations are also not so small that selection is typically overwhelmed by genetic drift [@Duthie2016a].

>>>>

Model
===================================================

@@

<!--- Note given previous comments: I've reworded because we don't know that pre-copulatory inbreeding avoidance is facilitating evolution of polyandry, not anymore. With the addition of post-copulatory strategy, it might be hindering it instead. If we wanted the answer to this, we would need to fix different levels of post-copulatory inbreeding avoidance and see if we get the same pattern as in Duthie et al. (2016) Evolution. We might not see the same result of pre-copulatory inbreeding avoidance facilitating evolution of polyandry. --->

<!--- The above point is especially problematic here because in the absence of post-copulatory inbreeding avoidance (in the current model) pre-copulatory inbreeding avoidance definitely does not facilitate evolution of costly polyandry -- this only happens under the severe constraints of our previous Evolution paper. Hence, it would have been a bit strange to suggest that post-copulatory inbreeding avoidance works `alongside' pre-copulatory inbreeding avoidance when we know that pre-copulatory inbreeding avoidance isn't doing anything. --->

To address whether or not evolution of post-copulatory inbreeding avoidance alongside pre-copulatory inbreeding avoidance can facilitate evolution of polyandry, we compare simulations in which polyandry and pre-copulatory and post-copulatory inbreeding avoidance can all evolve, with otherwise identical simulations in which post-copulatory inbreeding avoidance cannot evolve given four different costs of polyandry ($c_{P} = \{0, 0.0025, 0.005,  0.01\}$). To prevent post-copulatory inbreeding avoidance from evolving, we sever the connection from $F_{a}$ to $F_{p}$ such that all $F_{g}$ genotypes cause random fertilisation with respect to kinship, so $F_{a}$ alleles have no phenotypic effect.

To address how asymmetric costs associated with pre-copulatory and post-copulatory inbreeding avoidance and polyandry affect the long-term persistence of reproductive strategies, we quantify change in $M_{a}$ and $F_{a}$ over generations in simulations where the cost of pre-copulatory inbreeding strategy $c_{M}=0$ but the cost of post-copulatory inbreeding strategy $c_{F}=0.02$, and vice versa. We compare these simulations with evolution of a costly strategy in the absence of evolution of an alternative strategy (e.g., evolution of pre-copulatory inbreeding strategy when post-copulatory inbreeding strategy phenotype is fixed at zero, $F_{p}=0$). In Supporting Information, we present results for all possible cost combinations of $0$ and $0.02$.

To address how selection on an initially rare strategy of pre-copulatory or post-copulatory inbreeding avoidance is affected by the existence of the other strategy of inbreeding avoidance already being fixed, we first used exploratory simulations to quantify evolution of pre-copulatory inbreeding strategy, and post-copulatory inbreeding strategy and associated polyandry, in isolation. Then, to test whether pre-copulatory inbreeding avoidance would evolve when adaptive polyandry and post-copulatory inbreeding avoidance was already established, we initiated $M_{a}$ allele values at zero, but fixed $F_{a}$ allele values at $-10$ and $P_{a}$ allele values at $1$ (i.e., meaning that $F_{p}$ and $P_{p}$ did not evolve). Similarly, to test whether post-copulatory inbreeding avoidance would evolve when pre-copulatory inbreeding avoidance was already established, we initiated $F_{a}$ and $P_{a}$ allele values at zero but fixed $M_{a}$ allele values at $-10$. Consequently, because females have 10 diploid loci, when $M_{a}$ or $F_{a}$ alleles were fixed at $-10$, outbred females were $51$ times less likely to choose a full brother and $13.5$ times times less likely to choose a first cousin than a non-relative in pre-copulatory and post-copulatory choice, respectively.

In all simulations, we recorded mean values of $P_{a}$, $M_{a}$, and $F_{a}$ in each generation, and we present these values over generations to infer selection on phenotypes ($P_{p}$, $M_{p}$, and $F_{p}$). Each combination of parameter values simulated was replicated $40$ times, and grand mean values and standard errors of means are calculated in each generation across replicates. These analyses allow us to infer how allele values change over generations in response to costs, but also in response to the changing values of other alleles and therefore potential evolutionary feedbacks among reproductive strategies (e.g., selection on post-copulatory inbreeding strategy might be weak if polyandry is rare). For all replicates, we set the maximum number of generations to 40000, which exploratory simulations and previous modelling [@Duthie] showed to be sufficient for inferring long-term dynamics of mean allele values and therefore selection on phenotypes.  For all replicates, we set $\rho=5$ immigrants, which produced a range of kin and non-kin in each generation allowing females to express inbreeding strategies, and $n=8$ offspring, which was sufficient to keep populations consistently at carrying capacities and avoid population extinction. Values of $K_{f}$ and $K_{m}$ were both set to 100 because previous modelling showed that populations of this size are small enough that mate encounters between kin occur with sufficient frequency for selection on inbreeding strategy, but not so small that selection is typically overwhelmed by genetic drift [@Duthie2016a].

@@
